### 2020-06-16 TI_02_congruence.Rmd
next step: congruence measures

---
  title: "Triangulating iconicity: Using coding to predict experimental outcomes"
author: "Mark Dingemanse & Stella Punselie"
date: "Updated `r format(Sys.Date())`"
output: github_document
editor_options: 
  chunk_output_type: console
---

  ```{r global_options, include=FALSE}
knitr::opts_chunk$set(fig.width=8, fig.height=6, fig.path='out/',
                      echo=TRUE, warning=FALSE, message=FALSE)

```

Code notebook for a study of the relation between linguistically informed iconicity coding and experimentally collected guessability scores.

## Setup

```{r preliminaries, results="hide"}
# Packages
list.of.packages <- c("tidyverse","readxl","writexl")
new.packages <- list.of.packages[!(list.of.packages %in% installed.packages()[,"Package"])]
if(length(new.packages)) install.packages(new.packages)
lapply(list.of.packages, require, character.only=T)

# useful functions
`%notin%` <- function(x,y) !(x %in% y) 
mean.na <- function(x) mean(x, na.rm = T)
sd.na <- function(x) sd(x, na.rm = T)

```

## Data
Let's first load the original data and generate derivative XLS files to aid in the calibration.

```{r data}

d1 = read_excel("data\\ideophones_coded_MD_original.xlsx") %>% arrange(filename)
d2 = read_excel("data\\ideophones_coded_SP_original.xlsx") %>% arrange(filename)
d2 <- d2[,-c(1:7)] # remove non-unique fields except filename prior to merge

d <- merge(d1,d2,by="filename")

# sanity check: are all ideophones in d1 and d2 also in d? (0 = yes)
d1$filename[d1$filename %notin% d$filename]
d2$filename[d2$filename %notin% d$filename]

d.F <- d %>%
  dplyr::select(c("filename","ideophone"),starts_with("F_")) %>%
  dplyr::select(-starts_with("F_notes"),everything()) # put note fields at end
write_xlsx(d.F,path="data\\ideophones_F_consistency.xlsx")

d.M <- d %>%
  dplyr::select(c("filename","ideophone"),starts_with("M_")) %>%
  dplyr::select(-starts_with("M_notes"),everything()) # put note fields at end
write_xlsx(d.M,path="data\\ideophones_M_consistency.xlsx")

# function to get the names of coding categories (note: this presupposes there is a coder _MD, change as needed)
get_coding_categories <- function(data) {
  coder_1_names <- data %>% dplyr::select(ends_with("_MD")) %>% names()
  gsub('_MD','',coder_1_names)
}