### 2020-06-16 triangulating_iconicity.Rmd

---
title: "Triangulating iconicity"
author: "Mark Dingemanse & Stella Punselie"
date: "Updated `r format(Sys.Date())`"
output: github_document
editor_options: 
  chunk_output_type: console
---

```{r global_options, include=FALSE}
knitr::opts_chunk$set(fig.width=8, fig.height=6, fig.path='out/',
                      echo=TRUE, warning=FALSE, message=FALSE)

```

## Predicting iconicity
Intro

## Setup

```{r preliminaries, results="hide"}
# Clear workspace
rm(list=ls())
# Packages
list.of.packages <- c("ggplot2","dplyr","tidyr","readxl","ggthemes","lme4","viridis","stringdist","VGAM")
new.packages <- list.of.packages[!(list.of.packages %in% installed.packages()[,"Package"])]
if(length(new.packages)) install.packages(new.packages)
lapply(list.of.packages, require, character.only=T)

# gg_jitterbox
source("gg_jitterbox.R")

```

## Data

```{r data}

d1 = read_excel("data\\ideophones_for_coding_new_MD.xlsx")
d2 = read_excel("data\\ideophones_for_coding_new_SP.xlsx")

# add logodds (it's safer to predict against logodds than raw proportion correct)
d <- d %>%
  mutate(cumulative_congruence = FM_congruent) %>%
  group_by(paper) %>%
  mutate(logodds = probit(score))

# convert congruency measures to factors
tofactors <- "language|category|paper|group|FM_congruent"
d[,grep(tofactors,names(d))] <- lapply(d[,grep(tofactors,names(d))], as.factor)

```