### 2020-06-17 TI_01_consistency.Rmd
more work on congruency

---
title: "Triangulating iconicity: Coding, consistency & calibration"
author: "Mark Dingemanse & Stella Punselie"
date: "Updated `r format(Sys.Date())`"
output: github_document
editor_options: 
  chunk_output_type: console
---

```{r global_options, include=FALSE}
knitr::opts_chunk$set(fig.width=8, fig.height=6, fig.path='figures_md/',
                      echo=TRUE, warning=FALSE, message=FALSE)

```

## Coding consistency
Code notebook for combining coding from two coders, working out consistency, and recording coding/recoding decisions.

## Setup

```{r preliminaries, results="hide"}
# Packages
list.of.packages <- c("tidyverse","readxl","writexl","irr")
new.packages <- list.of.packages[!(list.of.packages %in% installed.packages()[,"Package"])]
if(length(new.packages)) install.packages(new.packages)
lapply(list.of.packages, require, character.only=T)

# useful functions
`%notin%` <- function(x,y) !(x %in% y) 
mean.na <- function(x) mean(x, na.rm = T)
sd.na <- function(x) sd(x, na.rm = T)

```

## Data
Let's first load the original data and generate derivative XLS files to aid in the calibration.

```{r data}

d1 = read_excel("data\\ideophones_coded_MD_original.xlsx") %>% arrange(filename)
d2 = read_excel("data\\ideophones_coded_SP_original.xlsx") %>% arrange(filename)
d2 <- d2[,-c(1:7)] # remove non-unique fields except filename prior to merge

d <- merge(d1,d2,by="filename")

d.F <- d %>%
  dplyr::select(c("filename","ideophone"),starts_with("F_")) %>%
  dplyr::select(-starts_with("F_notes"),everything()) # put note fields at end
#write_xlsx(d.F,path="data\\ideophones_F_consistency.xlsx")

d.M <- d %>%
  dplyr::select(c("filename","ideophone"),starts_with("M_")) %>%
  dplyr::select(-starts_with("M_notes"),everything()) # put note fields at end
#write_xlsx(d.M,path="data\\ideophones_M_consistency.xlsx")

# function to get the names of coding categories (note: this presupposes there is a coder _MD, change as needed)
get_coding_categories <- function(data) {
  coder_1_names <- data %>% dplyr::select(ends_with("_MD")) %>% names()
  gsub('MD','',coder_1_names)
}

```

## Consistency
We compute coding consistency for the original coding to establish which categories of the coding scheme are usable as is and which need recalibration or reformulation. We use `irr` package to compute % agreement and Cohen's Kappa.

```{r irr}

# build functions to make it less repetitive
get_agree <- function(data,cat) {
  # takes a dataset and a coding category, looks up fields of that coding category, and computes agreement
  agree(dplyr::select(data,starts_with(cat)))$value
}
get_kappa <- function(data,cat) {
  # takes a dataset and a coding category, looks up fields of that coding category, and computes kappa
  kappa2(dplyr::select(data,starts_with(cat)))$value
}

# sanity check: does the function work? yes
get_agree(d.F,cat='F_redup_')

# get list of coding categories and create dataframes for the consistency values
F_categories <- get_coding_categories(d.F)
F_consistency <- as.data.frame(F_categories) %>% dplyr::rename(category = F_categories)
M_categories <- get_coding_categories(d.M)
M_consistency <- as.data.frame(M_categories) %>% dplyr::rename(category = M_categories)

F_consistency$agree <- NA
F_consistency$kappa <- NA
for (i in 1:(length(F_categories)-1)) { # -1 to exclude the notes column
  F_consistency$agree[i] <- (get_agree(data=d.F,cat=F_categories[i]))
  F_consistency$kappa[i] <- (get_kappa(data=d.F,cat=F_categories[i]))
}
F_consistency
```

For Form, mean agreement in the original coding is **`r round(mean.na(F_consistency$agree))`%** and the lowest Kappa values are for F_intonation, F_voice, and F_partialredup, pointing to the need to resolve inconsistencies and sharpen those coding categories.

```{r irr_M}
M_consistency$agree <- NA
M_consistency$kappa <- NA
for (i in 1:(length(M_categories)-1)) {
  M_consistency$agree[i] <- (get_agree(data=d.M,cat=M_categories[i]))
  M_consistency$kappa[i] <- (get_kappa(data=d.M,cat=M_categories[i]))
}
M_consistency

```

For Meaning, mean agreement in the original coding is **`r round(mean.na(M_consistency$agree))`%** and the lowest Kappa values are for M-durative, M_weight, and M_irregular, pointing to the need to resolve inconsistencies and sharpen those coding categories.

## Outcomes
Using the consistency scores as a guide, we revisited the most problematic categories to (i) recalibrate coders' understanding of the categories, (ii) reformulate instructions and (iii) in some cases rename the coding category to reflect the revisions.

### Form
1.	*F_redupmod* was renamed (originally F_partialredup) and the coding instructions were sharpened to reflect the fact that we want to capture both partial and modified forms of reduplication.
2. *F_voice*: coding instructions were clarified to note that this question privileges consonants that crosslinguistically often enter into voicing oppositions (plosives and fricatives) as well as nasals, documented for their iconic uses across languages. This is close to Westermann’s use of this feature. We also considered a more continuous version of this feature that would weight or classify consonants based on their place on the sonority scale, but decided against it because the coding was already too fine-grained.
3. *F_finallength* was renamed (originally F_finalvowel) to capture word final lengthening of any final sound, whether final vowel or nasal, glide or trill. 
4.	*F_intonation*: coding instructions were clarified to include a decision procedure for forms with falling or rising intonation.

### Meaning
1. *M_irregularity*: coding instructions were modified to indicate that the primary semantic property here is one of irregularity in the sense of imbalance (a highly regular ridged pattern would not qualify)
2.	*M_weight*: coding instructions were sharpened to focus on binary oppositions of size, weight and brightness and to exclude Westermann’s hard/soft distinction, which we found orthogonal and conflicting in some cases. This also means that the middle value essentially means neutral or “not applicable”. It may be advisable to further break this up this feature, as there is no guarantee that size, weight and brightness pattern similarly across languages.
3. *M_durative*: coding instructions were sharpened to clarify that this targets events, not states and to clarify its relation to M_distribution.

The original coding sheet is available as X and the updated coding sheet as X_version_2.

## Calibrated coding

After reformulating and revising the coding scheme, we discussed all inconsistent cases to reach a consensus and recoded accordingly. Here we check whether we now indeed have a canonical, ground truth set of coded data.

```{r consistency_2ndround}
d1 = read_excel("data\\ideophones_coded_MD.xlsx") %>% arrange(filename)
d2 = read_excel("data\\ideophones_coded_SP.xlsx") %>% arrange(filename)
d2 <- d2[,-c(1:7)] # remove non-unique fields except filename prior to merge

d <- merge(d1,d2,by="filename")

d.F <- d %>%
  dplyr::select(c("filename","ideophone","meaning","meaning_NL"),starts_with("F_")) %>%
  dplyr::select(-starts_with("F_notes"),everything()) # put note fields at end

d.M <- d %>%
  dplyr::select(c("filename","ideophone","meaning","meaning_NL"),starts_with("M_")) %>%
  dplyr::select(-starts_with("M_notes"),everything()) # put note fields at end

# get list of coding categories and create dataframes for the consistency values
F_categories <- get_coding_categories(d.F)
F_consistency <- as.data.frame(F_categories) %>% dplyr::rename(category = F_categories)
M_categories <- get_coding_categories(d.M)
M_consistency <- as.data.frame(M_categories) %>% dplyr::rename(category = M_categories)

F_consistency$agree <- NA
F_consistency$kappa <- NA
for (i in 1:(length(F_categories)-1)) { # -1 to exclude the notes column
  F_consistency$agree[i] <- (get_agree(data=d.F,cat=F_categories[i]))
  F_consistency$kappa[i] <- (get_kappa(data=d.F,cat=F_categories[i]))
}
F_consistency

```

Hmm, a consistency check at this point shows we're nearly there but we're not 100% consistent on F_redup, F_monosyllabic and F_vowelquality and F_finallength. We can export only those columns to have a quick check.

```{r inconsistency_output}

inconsistent <- c('F_redup','F_monosyllabic','F_vowelquality','F_finallength')
d.F.inconsistent <- d.F %>%
  dplyr::select(filename,ideophone,meaning,meaning_NL,matches(paste(inconsistent,collapse="|"))) %>%
  mutate(F_redup_C = ifelse(F_redup_MD == F_redup_SP,1,0),
         F_monosyllabic_C = ifelse(F_monosyllabic_MD == F_monosyllabic_SP,1,0),
         F_vowelquality_C = ifelse(F_vowelquality_MD == F_vowelquality_SP,1,0),
         F_finallength_C = ifelse(F_finallength_MD == F_finallength_SP,1,0)) %>%
  mutate(check = F_redup_C + F_monosyllabic_C + F_vowelquality_C + F_finallength_C) %>%
  filter(check < 4)
d.F.inconsistent <- d.F.inconsistent[, order(names(d.F.inconsistent))]

write_xlsx(d.F.inconsistent,path="data\\ideophones_F_consistency_2ndround.xlsx")

```

* Ultimately we want to output a consensus or ground truth file `ideophones_coded.xlsx`, to be used in the next step of the analyses.

```{r write_final}

# for now I'm taking just one coder
d.coded <- read_excel("data\\ideophones_coded_SP.xlsx") %>% 
  arrange(filename) %>%
  dplyr::select(-matches('notes'))
names(d.coded) <- gsub('_SP','',names(d.coded))

notes <- d %>% dplyr::select(filename,matches('notes'))

d.coded <- left_join(d.coded,notes)

write_xlsx(d.coded,path="data\\ideophones_coded.xlsx")

```