### 2020-06-17 TI_02_congruence.Rmd
update

---
title: "Triangulating iconicity: From structure mapping to guessability"
author: "Mark Dingemanse & Stella Punselie"
date: "Updated `r format(Sys.Date())`"
output: github_document
editor_options: 
  chunk_output_type: console
---

  ```{r global_options, include=FALSE}
knitr::opts_chunk$set(fig.width=8, fig.height=6, fig.path='figures_md/',
                      echo=TRUE, warning=FALSE, message=FALSE)

```

Code notebook for a study of the relation between linguistically informed iconicity coding and experimentally collected guessability scores.

## Setup

```{r preliminaries, results="hide"}
# Packages
list.of.packages <- c("tidyverse","readxl","writexl","ggthemes","viridis","lme4","VGAM")
new.packages <- list.of.packages[!(list.of.packages %in% installed.packages()[,"Package"])]
if(length(new.packages)) install.packages(new.packages)
lapply(list.of.packages, require, character.only=T)

# useful functions
`%notin%` <- function(x,y) !(x %in% y) 
mean.na <- function(x) mean(x, na.rm = T)
sd.na <- function(x) sd(x, na.rm = T)

# custom functions
source("TI_functions.R")
```

## Data
Here we load the ground truth version of the coded data and add the independently collected guessability scores from the 2016 Collabra and Language papers.

```{r data}

# get coding data
d = read_excel("data\\ideophones_coded.xlsx") %>% arrange(filename)

coding_categories <- d %>% dplyr::select(matches('_'),-matches('notes|meaning')) %>% names()

# add guessability scores from the Collabra and Language studies.

d.scores = read_excel("data\\ideophones_guessability.xlsx") %>%
  dplyr::select(-category)
d <- left_join(d,d.scores,by=c("ideophone","language","study" = "paper"))

# add logodds (for when we run stats: it's more sensible to predict against logodds than raw proportion correct)
d <- d %>%
  group_by(study) %>%
  mutate(logodds = probitlink(score))

```

We add congruency measures that link F and M elements.

* TO DISCUSS: We need to think about the evidential value of each of the congruency measures. Most apply to many words, which is good. If some of them only capture a few (as `C_durative` and `C_weight_tone` seem to do), they are not so interesting because they don't really capture recurring iconic mappings in the data. Or perhaps this may indicate a need to recode — e.g. perhaps `F_finallength` should not just map onto aspectual semantics (like `M_durative` but also to shape, sth like `M_long`: a very concrete kind of mapping.)

```{r congruency_1}

# using ifelse() statements that essentially say, if these conditions are true, use 1, otherwise 0
d <- d %>%
  mutate(C_modality = ifelse(M_sound == 1,1,0),
         C_iterative = ifelse(F_redup == 1 & M_distribution == 1,1,0),
         C_irregular = ifelse(F_redupmod == 1 & M_irregular == 1,1,0),
         C_closure = ifelse(F_closedsyllable == 1 & M_abrupt == 1,1,0),
         C_punctual = ifelse(F_monosyllabic == 1 & M_punctual == 1,1,0),
         C_durative = ifelse(F_finallength == 1 & M_durative == 1,1,0),
         C_weight_voice = ifelse((F_voice == 0 & M_weight == 0) | (F_voice == 2 & M_weight) == 2,1,0),
         C_weight_vowel = ifelse((F_vowelquality == 0 & M_weight == 0) | (F_vowelquality == 2 & M_weight == 2),1,0),
         C_weight_tone = ifelse((F_intonation == 0 & M_weight == 2) | (F_intonation == 2 & M_weight == 0),1,0)
         ) %>%
  mutate(C_cumulative = C_modality + C_iterative + C_irregular + C_closure + C_punctual + C_durative + C_weight_voice + C_weight_vowel + C_weight_tone) %>%
  mutate(C_any = ifelse(C_cumulative == 0,0,1),
         C_atleast2 = ifelse(C_cumulative > 1,1,0),
         C_ternary = ifelse(C_cumulative == 0,0,
                            ifelse(C_cumulative == 1,1,2)))

d %>%
  group_by(C_cumulative) %>%
    summarise(n=n())

```

## Plots

A quick first look at some congruency measures. Uses a custom plotting function `icoplot()`, which is a wrapper for a ggplot object constructed with `geom_dotplot`.

```{r icoplots}

# make a copy dp of the data in which we convert some measures to factors for easy plotting & facetting
tofactors <- paste(c("language|category|study|group",names(d[grep('C_',names(d))])),collapse = "|")
dp <- d
dp[,grep(tofactors,names(dp))] <- lapply(dp[,grep(tofactors,names(dp))], as.factor)

# all data points by language
icoplot()

icoplot("C_modality")
icoplot("C_iterative")
icoplot("C_irregular")
icoplot("C_punctual")
icoplot("C_durative")
icoplot("C_weight_voice")
icoplot("C_weight_vowel")
icoplot("C_weight_tone")

icoplot("C_cumulative")

icoplot("C_any")
icoplot("C_atleast2")
icoplot("C_ternary")

# blank plot can be useful as background for slides
ggplot(data=dp, aes(x=study,y=score)) +
  theme_tufte() + ylim(0,1) + theme(legend.position="none",axis.title.x=element_blank(),plot.margin=margin(0,0,10,0)) + 
  stat_summary(fun.y=median,geom="point",size=8,shape=21,stroke=1,fill="white",colour="#c9c9c9") +
  geom_dotplot(colour="white",fill="#c9c9c9",stackgroups=T,dotsize=1.5,binwidth=0.01,binaxis="y",stackdir = "center")

```

## Exploring congruency blind spots
What are we not capturing? Our coding and congruency measures are designed to be reproducible and human-codable. They are not exhaustive, and there may well be aspects that we are missing. Here we explore the congruency data to find highly guessable ideophones that our coding did not (yet) identify as having iconic mappings.

```{r congruency_blindspots}

d.blindspots <- d %>%
  filter(C_cumulative == 0, score >0.5) %>%
  dplyr::select(filename,ideophone,meaning,meaning_NL,language,score) %>%
  arrange(-score)
View(d.blindspots)

```

## Stats

```{r congruency_stats_1}

# simplest thing we can do is a correlation

cor.test(d$logodds,d$C_cumulative)
cor.test(d$score,d$C_cumulative)

```