### 2020-06-26 TI_03_ratings.Rmd

---
  title: "Triangulating iconicity: Iconicity ratings"
author: "Mark Dingemanse & Stella Punselie"
date: "Updated `r format(Sys.Date())`"
output: github_document
editor_options: 
  chunk_output_type: console
---

  ```{r global_options, include=FALSE}
knitr::opts_chunk$set(fig.width=8, fig.height=6, fig.path='figures_md/',
                      echo=TRUE, warning=FALSE, message=FALSE)

```

Code notebook for a study of the relation between iconicity ratings and experimentally collected guessability scores.

## Setup

```{r preliminaries, results="hide"}
# Packages
list.of.packages <- c("tidyverse","readxl","writexl","ggthemes","viridis","lme4","VGAM")
new.packages <- list.of.packages[!(list.of.packages %in% installed.packages()[,"Package"])]
if(length(new.packages)) install.packages(new.packages)
lapply(list.of.packages, require, character.only=T)

# useful functions
`%notin%` <- function(x,y) !(x %in% y) 
mean.na <- function(x) mean(x, na.rm = T)
sd.na <- function(x) sd(x, na.rm = T)

```

## Data
Here we load the combined coding and guessability data and then add the rating data.

```{r data}
# get coding data
d = read_excel("data\\ideophones_coded.xlsx") %>% arrange(filename)

coding_categories <- d %>% dplyr::select(matches('_'),-matches('notes|meaning')) %>% names()

# add guessability scores from the Collabra and Language studies.

d.scores = read_excel("data\\ideophones_guessability.xlsx") %>%
  dplyr::select(-category)
d <- left_join(d,d.scores,by=c("ideophone","language","study" = "paper"))

# add logodds (for when we run stats: it's more sensible to predict against logodds than raw proportion correct)
d <- d %>%
  group_by(study) %>%
  mutate(logodds = probitlink(score))

# add mean ratings
d.ratings = read_excel("data\\ideophones_rated_means.xlsx") %>%
  dplyr::select(-c(item,category,list))
d <- left_join(d,d.ratings,by=c("filename","language","study"))

# convert ratings to z-scores for comparison with guessability (logodd) scores
d$rating_z <- scale(d$rating,center=T,scale=T)

```

## Statistics
```{r stats}
# before calculating the correlation coefficient, we visualise the correlation between the guessing and rating data
ggplot(d,aes(x=rating_z,y=logodds)) + geom_point(colour="dark grey",position="jitter") + geom_smooth(method=lm,colour="black")
ggsave("figures\\TI_03_corr.png",height=5,width=7.5)

ggplot(d,aes(x=rating_z,y=logodds)) + geom_point(aes(color=category,shape=language),position="jitter")
ggsave("figures\\TI_03_corr_bylanguageandcategory.png",height=5,width=7.5)

# Pearson correlation
cor.test(d$logodds,d$rating_z,method="pearson")

```

Pearson's correlation coefficient indicates that there is a strong positive correlation between the guessing and rating scores, r(237) = .57, p < .001.