### 2024-04-03 legacy/TI_04_triangulation.Rmd

---
  title: "Triangulating iconicity: combining coding, guessing and rating measures"
author: "Mark Dingemanse & Stella Punselie"
date: "Updated `r format(Sys.Date())`"
output: github_document
editor_options: 
  chunk_output_type: console
---

  ```{r global_options, include=FALSE}
knitr::opts_chunk$set(fig.width=8, fig.height=6, fig.path='figures_md/',
                      echo=TRUE, warning=FALSE, message=FALSE)

```

Code notebook for a study of the relation between linguistically informed iconicity coding and experimentally collected guessability scores.

## Setup

```{r preliminaries, results="hide"}
# Packages
list.of.packages <- c("tidyverse","readxl","writexl","ggthemes","gghalves","plotly","viridis")
new.packages <- list.of.packages[!(list.of.packages %in% installed.packages()[,"Package"])]
if(length(new.packages)) install.packages(new.packages)
lapply(list.of.packages, require, character.only=T)

# useful functions
`%notin%` <- function(x,y) !(x %in% y) 
mean.na <- function(x) mean(x, na.rm = T)
sd.na <- function(x) sd(x, na.rm = T)

```
## Data

```{r data}
d = read.csv("data\\ideophones_coded_guessed_rated.csv")

coding_categories <- d %>% dplyr::select(matches('F_|M_'),-matches('notes|meaning')) %>% names()

# add cumulative score to allow summing of cumulative, rating, guessability
d$cumulative_z <- scale(d$C_cumulative,center=T,scale=T)
d <- d %>%
  mutate(overallrank = as.numeric(cumulative_z + rating_z + score_z))

d %>% dplyr::select(ideophone,
                    meaning,
                    meaning_NL,
                    category,
                    language,
                    study,
                    score,
                    rating,
                    C_cumulative,
                    overallrank) %>%
  filter(score > .65,
         category == "Motion",
         C_cumulative > 0) %>%
  dplyr::arrange(-overallrank) %>%
  View()

# make a copy dp of the data in which we convert some measures to factors for easy plotting & facetting
tofactors <- paste(c("language|category|study|group",names(d[grep('C_',names(d))])),collapse = "|")
dp <- d
dp[,grep(tofactors,names(dp))] <- lapply(dp[,grep(tofactors,names(dp))], as.factor)

```

## Visualizations
Let's have a look at the relations between the coding, guessing and rating data.
```{r visuals}

ggplot(d,aes(x=rating_z,y=logodds, color=C_cumulative)) +
  geom_point() +
  geom_smooth(method=loess,color="black") +
  scale_fill_viridis(option="plasma",discrete=F,begin=0.3,end=0.9) +
  scale_colour_viridis(option="plasma",discrete=F,begin=0.3,end=0.9) +
  NULL

d %>%
  plot_ly(x = ~rating_z, 
          y = ~logodds, 
          z = ~C_cumulative,
          type="scatter3d",
          mode="markers",
          color= ~rating_z,
          name=~ideophone,
          showlegend=F) %>%
  layout(showlegend=F)

d %>%
  plot_ly(x = ~rating_z, 
          y = ~logodds, 
          type= "scatter",
          color= ~rating_z,
          name=~ideophone,
          showlegend=F) %>%
  layout(showlegend=F)

ggplot(dp, aes(x=C_cumulative,y=rating_z,fill=score_z,color=score_z)) +
  theme_tufte() +
  geom_half_boxplot() +
  geom_half_point() +
  NULL

## look at some data

d %>%
  filter(rating_z >1,
         logodds >0.6,
         C_cumulative >1) %>%
  View()

```