### 2024-04-03 paper_01_consistency.Rmd
use native pipe & check whether everything still renders

---
title: "Coding, consistency & calibration"
author: "[anonymized for review]"
date: "Updated `r format(Sys.Date())`"
output: github_document
editor_options: 
  chunk_output_type: console
---

```{r global_options, include=FALSE}
knitr::opts_chunk$set(fig.width=8, fig.height=6, fig.path='figures_md/',
                      echo=TRUE, warning=FALSE, message=FALSE)

```

## Coding consistency
Code notebook for combining coding from two coders, working out consistency, and recording coding/recoding decisions.

## Setup

```{r preliminaries, results="hide"}
# Packages
list.of.packages <- c("tidyverse","readxl","writexl","irr")
# new.packages <- list.of.packages[!(list.of.packages %in% installed.packages()[,"Package"])]
# if(length(new.packages)) install.packages(new.packages)
lapply(list.of.packages, require, character.only=T)

# useful functions
`%notin%` <- function(x,y) !(x %in% y) 
mean.na <- function(x) mean(x, na.rm = T)
sd.na <- function(x) sd(x, na.rm = T)

```

## Data
Let's first load the original data and generate derivative XLS files to aid in the calibration.

```{r data}

d1 = read_excel("data/ideophones_coded_MD_original.xlsx") |> arrange(filename)
d2 = read_excel("data/ideophones_coded_SP_original.xlsx") |> arrange(filename)
d2 <- d2[,-c(1:7)] # remove non-unique fields except filename prior to merge

d <- merge(d1,d2,by="filename")

d.F <- d |>
  dplyr::select(c("filename","ideophone"),starts_with("F_")) |>
  dplyr::select(-starts_with("F_notes"),everything()) # put note fields at end
#write_xlsx(d.F,path="data\\ideophones_F_consistency.xlsx")

d.M <- d |>
  dplyr::select(c("filename","ideophone"),starts_with("M_")) |>
  dplyr::select(-starts_with("M_notes"),everything()) # put note fields at end
#write_xlsx(d.M,path="data\\ideophones_M_consistency.xlsx")

# function to get the names of coding categories (note: this presupposes there is a coder _MD, change as needed)
get_coding_categories <- function(data) {
  coder_1_names <- data |> dplyr::select(ends_with("_MD")) |> names()
  gsub('MD','',coder_1_names)
}

```

## Consistency
We compute coding consistency for the original coding to establish which categories of the coding scheme are usable as is and which need recalibration or reformulation. We use `irr` package to compute % agreement and Cohen's Kappa.

```{r irr}

# build functions to make it less repetitive
get_agree <- function(data,cat) {
  # takes a dataset and a coding category, looks up fields of that coding category, and computes agreement
  agree(dplyr::select(data,starts_with(cat)))$value
}
get_kappa <- function(data,cat) {
  # takes a dataset and a coding category, looks up fields of that coding category, and computes kappa
  kappa2(dplyr::select(data,starts_with(cat)))$value
}

# sanity check: does the function work? yes
get_agree(d.F,cat='F_redup_')

# get list of coding categories and create dataframes for the consistency values
F_categories <- get_coding_categories(d.F)
F_consistency <- as.data.frame(F_categories) |> dplyr::rename(category = F_categories)
M_categories <- get_coding_categories(d.M)
M_consistency <- as.data.frame(M_categories) |> dplyr::rename(category = M_categories)

F_consistency$agree <- NA
F_consistency$kappa <- NA
for (i in 1:(length(F_categories)-1)) { 
  F_consistency$agree[i] <- (get_agree(data=d.F,cat=F_categories[i]))
  F_consistency$kappa[i] <- (get_kappa(data=d.F,cat=F_categories[i]))
}
F_consistency
mean.na(F_consistency$kappa)

```

For Form, mean agreement in the original coding is **`r round(mean.na(F_consistency$agree))`%** and the lowest Kappa values are for F_intonation, F_voice, and F_partialredup, pointing to the need to resolve inconsistencies and sharpen those coding categories.

```{r irr_M}
M_consistency$agree <- NA
M_consistency$kappa <- NA
for (i in 1:(length(M_categories)-1)) {
  M_consistency$agree[i] <- (get_agree(data=d.M,cat=M_categories[i]))
  M_consistency$kappa[i] <- (get_kappa(data=d.M,cat=M_categories[i]))
}
M_consistency
mean.na(M_consistency$agree)

```

For Meaning, mean agreement in the original coding is **`r round(mean.na(M_consistency$agree))`%** and the lowest Kappa values are for M_durative, M_weight, and M_irregular, pointing to the need to resolve inconsistencies and sharpen those coding categories.

## Outcomes
Using the consistency scores as a guide, we revisited the most problematic categories to (i) recalibrate coders' understanding of the categories, (ii) reformulate instructions and (iii) in some cases rename the coding category to reflect the revisions.

### Form
1.	*F_redupmod* was renamed (originally F_partialredup) and the coding instructions were sharpened to reflect the fact that we want to capture both partial and modified forms of reduplication.
2. *F_voice*: coding instructions were clarified to note that this question privileges consonants that crosslinguistically often enter into voicing oppositions (plosives and fricatives) as well as nasals, documented for their iconic uses across languages. This is close to Westermann’s use of this feature. We also considered a more continuous version of this feature that would weight or classify consonants based on their place on the sonority scale, but decided against it because the coding was already too fine-grained.
3. *F_finallength* was renamed (originally F_finalvowel) to capture word final lengthening of any final sound, whether final vowel or nasal, glide or trill. 
4.	*F_intonation*: coding instructions were clarified to include a decision procedure for forms with falling or rising intonation.

### Meaning
1. *M_irregularity*: coding instructions were modified to indicate that the primary semantic property here is one of irregularity in the sense of imbalance (a highly regular ridged pattern would not qualify)
2.	*M_weight*: coding instructions were sharpened to focus on binary oppositions of size, weight and brightness and to exclude Westermann’s hard/soft distinction, which we found orthogonal and conflicting in some cases. This also means that the middle value essentially means neutral or “not applicable”. It may be advisable to further break this up this feature, as there is no guarantee that size, weight and brightness pattern similarly across languages.
3. *M_long* was renamed (originally M_durative): coding instructions were broadened to capture percepts that are long in time and/or space (in line with M_distribution), and sharpened to clarify that for the temporal dimension it targets only events.

The original coding sheet is available as `predico_coding_scheme_1.pdf` and the updated coding sheet as `predico_coding_scheme_2.pdf`.

## Calibrated coding

After reformulating and revising the coding scheme, we discussed all inconsistent cases to reach a consensus and recoded accordingly. Here we check whether we now indeed have a canonical, ground truth set of coded data.

```{r consistency_2ndround}
d1 = read_excel("data/ideophones_coded_MD.xlsx") |> arrange(filename)
d2 = read_excel("data/ideophones_coded_SP.xlsx") |> arrange(filename)
d2 <- d2[,-c(1:7)] # remove non-unique fields except filename prior to merge

d <- merge(d1,d2,by="filename")

d.F <- d |>
  dplyr::select(c("filename","ideophone","meaning","meaning_NL"),starts_with("F_")) |>
  dplyr::select(-starts_with("F_notes"),everything()) # put note fields at end

d.M <- d |>
  dplyr::select(c("filename","ideophone","meaning","meaning_NL"),starts_with("M_")) |>
  dplyr::select(-starts_with("M_notes"),everything()) # put note fields at end

# get list of coding categories and create dataframes for the consistency values
F_categories <- get_coding_categories(d.F)
F_consistency <- as.data.frame(F_categories) |> dplyr::rename(category = F_categories)
M_categories <- get_coding_categories(d.M)
M_consistency <- as.data.frame(M_categories) |> dplyr::rename(category = M_categories)

F_consistency$agree <- NA
F_consistency$kappa <- NA
for (i in 1:(length(F_categories)-1)) { # -1 to exclude the notes column
  F_consistency$agree[i] <- (get_agree(data=d.F,cat=F_categories[i]))
  F_consistency$kappa[i] <- (get_kappa(data=d.F,cat=F_categories[i]))
}
F_consistency

```

A consistency check at this point shows we're nearly there but we're not 100% consistent on F_redup, F_monosyllabic and F_vowelquality and F_finallength. We can export only those columns to have a quick check. However, as we do not aim for 100% consistency, this is more for informative purposes.

```{r inconsistency_output}

inconsistent <- c('F_redup','F_monosyllabic','F_vowelquality','F_finallength')
d.F.inconsistent <- d.F |>
  dplyr::select(filename,ideophone,meaning,meaning_NL,matches(paste(inconsistent,collapse="|"))) |>
  mutate(F_redup_C = ifelse(F_redup_MD == F_redup_SP,1,0),
         F_monosyllabic_C = ifelse(F_monosyllabic_MD == F_monosyllabic_SP,1,0),
         F_vowelquality_C = ifelse(F_vowelquality_MD == F_vowelquality_SP,1,0),
         F_finallength_C = ifelse(F_finallength_MD == F_finallength_SP,1,0)) |>
  mutate(check = F_redup_C + F_monosyllabic_C + F_vowelquality_C + F_finallength_C) |>
  filter(check < 4)
d.F.inconsistent <- d.F.inconsistent[, order(names(d.F.inconsistent))]

write_xlsx(d.F.inconsistent,path="data/ideophones_F_consistency_2ndround.xlsx")

```

### Consensus version
The final result of the coding consistency is a consensus or ground truth file `ideophones_coded.xlsx`, to be used in the next step of the analyses. We do this this by taking the (calibrated and revised) data from one coder.

```{r create_consensus_version}

d.coded <- read_excel("data/ideophones_coded_SP.xlsx") |> 
  arrange(filename) |>
  dplyr::select(-matches('notes'))
names(d.coded) <- gsub('_SP','',names(d.coded))

notes <- d |> dplyr::select(filename,matches('notes'))

d.coded <- left_join(d.coded,notes)

```

## Adding congruency measures
We add congruency measures by combining F and M features as described in the coding scheme.

```{r congruency_1}

# note that we overwrite the old d with the new consensus version here

# using ifelse() statements that essentially say, if these conditions are true, use 1, otherwise 0
d <- d.coded |>
  mutate(C_modality = ifelse(M_sound == 1,1,0),
         C_iterative = ifelse(F_redup == 1 & M_distribution == 1,1,0),
         C_irregular = ifelse(F_redupmod == 1 & M_irregular == 1,1,0),
         C_closure = ifelse(F_closedsyllable == 1 & M_abrupt == 1,1,0),
         C_punctual = ifelse(F_monosyllabic == 1 & M_punctual == 1,1,0),
         C_long = ifelse(F_finallength == 1 & M_long == 1,1,0),
         C_weight_voice = ifelse((F_voice == 0 & M_weight == 0) | (F_voice == 2 & M_weight == 2),1,0),
         C_weight_vowel = ifelse((F_vowelquality == 0 & M_weight == 0) | (F_vowelquality == 2 & M_weight == 2),1,0),
         C_weight_tone = ifelse((F_intonation == 0 & M_weight == 2) | (F_intonation == 2 & M_weight == 0),1,0)
         )

# some mappings are mutually exclusive, so we combine them

#sanity check: are they indeed mutually exclusive? these calls should come up empty
d |> 
  mutate(C_aspect = C_iterative + C_punctual) |>
  dplyr::select(ideophone,meaning_NL,language,C_iterative,C_punctual,C_aspect) |>
  filter(C_aspect > 1)
d |>
  mutate(C_length = C_long + C_closure) |>
  dplyr::select(ideophone,meaning_NL,language,C_iterative,C_closure,C_length) |>
  filter(C_length > 1)

d <- d |>
  mutate(C_aspect = C_iterative + C_punctual,
         C_length = C_long + C_closure)

# C_weight_ subfeatures are linked to one and the same meaning feature M_weight so they are recombined into C_magnitude
# (note: this may be subject to revision, as the subfeatures may not deserve equal weight)

d <- d |>
  mutate(C_weight = C_weight_voice + C_weight_vowel + C_weight_tone,
         C_magnitude = ifelse(C_weight == 0,0,1)) # i.e. a magnitude value that is 1 if any weight feature was congruent

# add cumulative iconicity scores (and some others)
# any1 = 1 if an ideophone has at least 1 congruency feature of any type
# atleast2 = if an ideophone has at least 2 congruency features of any type
# ternary = simple way of conflating the (smaller) upper regions: 1=1, 2=2, >2 = 3
d <- d |>
  mutate(C_cumulative = C_modality + C_aspect + C_length + C_irregular + C_magnitude) |>
  mutate(C_any1 = ifelse(C_cumulative == 0,0,1),
         C_atleast2 = ifelse(C_cumulative > 1,1,0),
         C_ternary = ifelse(C_cumulative == 0,0,
                            ifelse(C_cumulative == 1,1,
                                   ifelse(C_cumulative == 2,2,3)))) # i.e. 1, 2 or 3

```

Finally, we write this to XLSX and CSV files. (Commented out here to avoid overwriting anytime we knit the markdown.)

```{r write_files}

#write_xlsx(d,path="data\\ideophones_coded.xlsx")
#write.csv(d,file="data\\ideophones_coded.csv")

```